\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-09-04 phd-cs-us.tex
Updates from Overleaf

@@

@@

It's important to aim for the top venues in your field. You can
find top CS conferences at places such as CSRankings~\cite{csrankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences}. Local conferences and non-English journals or conferences do
not carry as much weight since their quality is often unknown to US faculty.

@@

For Vietnamese students, it's worth highlighting the \href{https://vef2.org/}{VEF 2.0 program}, which has played a crucial role in aiding many good students in gaining acceptance to top-tier PhD programs in the US. VEF 2.0 follows an intriguing model where  US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf. Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters is remarkably positive. This fact is underscored by the noteworthy success of numerous VEF 2.0 students in securing admission to top CS PhD programs across the US.

@@

\begin{commentbox}[Vu]
I always read the research statement (and the LORs) first. If I am
persuaded by then, I would skim over other factors and advocate for
admission (unless I see red flags in other parts). If I am not
convinced, then I will likely recommend rejection (unless I see
something standout in other parts).
\tcblower
Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty. 
Also, be careful not to send statements to wrong schools or mixing
facts (e.g., talking about school X but mentioned about working with
profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
should.
\end{commentbox}

@@

@@

@@

\section{Miscs and FAQs}

@@

International students not familiar with US universities often put them into \emph{two} bins:  (i) very top schools that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.  Sometimes they rank CS programs using the reputations non-CS programs such as medical, math, or physics.
In some cases they rank universities based on popular states they know in the US, e.g., California and New York.  Let's just say there are so many thing wrong with these methods.

@@

      \begin{center}
        \includegraphics[scale=0.3]{c8.png}
      \end{center}

@@

Below is a good example that I would definitely reply to.

@@

\begin{commentbox}[Vu]
I've been called Dr. Vu and I find it a bit amusing but am totally fine with it.
\end{commentbox}

@@

\section{Virtual Research Opportunities Beyond Physical Boundaries}
\didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate. For this reason they often find themselves unprepared for PhD admission process by the time they finish undergraduate and have to spend a year or more to improve their application profiles. I think it would be beneficial to expand the document to also target students who are not yet ready for the application process to help them improve on all criteria considered during admission. I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier Ph.D. programs. However, students from underrepresented small colleges, both in the United States and internationally, often encounter limited research opportunities within their institutions. Thankfully, there exists an alternative pathway for these aspiring researchers to gain valuable research experience, even without being physically present at a university.

>>>>

\epigraph{"How is education supposed to make me feel smarter? Besides, every time I learn something new, it pushes some of the old stuff out of my brain. Remember when I took that home wine making course, and I forgot how to drive?"}{The Simpsons}

@@

\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{The Simpsons}

@@

It's important to aim for the top venues in your field. You can
find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences}. Local conferences and non-English journals or conferences do
not carry as much weight since their quality is often unknown to US faculty.

@@

For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF 2.0 program}, which has helped many good students in gaining acceptance to top-tier PhD programs in the US. VEF 2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf. Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters can be very positive as they have helped numerous VEF 2.0 students getting into top CS PhD programs in the US.

@@

\begin{commentbox}[Vu]
I often read the research statement (and the LORs) first. If I am
persuaded by then, I would skim over other factors and advocate for
admission (unless I see red flags in other parts). If I am not
convinced, then I will likely recommend rejection (unless I see
something standout in other parts).
\tcblower
Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty. 
Also, be careful not to send statements to wrong schools or mixing
facts (e.g., talking about school X but mentioned about working with
profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
should.
\end{commentbox}

@@

\epigraph{"Oh... and how is education supposed to make me feel smarter? Besides, every time I learn something new, it pushes some old stuff out of my brain. Remember when I took that home wine-making course and I forgot how to drive?"}{The Simpsons}

@@

\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{The Simpson}

@@

\section{Miscs and FAQs}
\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{The Simpsons}

@@

International students not familiar with US universities often put them into \emph{two} bins:  (i) very top schools that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.  In many cases, they use resources such as US News, but their ranking is \emph{highly} questionable~\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.  Sometimes these students rank CS programs using the reputations non-CS programs such as medical, math, or physics.
In some cases they rank universities based on popular states they know in the US, e.g., California and New York.  Clearly, there are so many thing wrong with these methods.

@@

      \begin{center}
        \includegraphics[scale=0.4]{c8.png}
      \end{center}

@@

Below is a good example that I would reply to.

@@

\begin{commentbox}[Vu]
Dr. Vu is a bit amusing but I am totally fine with it.
\end{commentbox}

@@

\section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
\didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate. For this reason they often find themselves unprepared for PhD admission process by the time they finish undergraduate and have to spend a year or more to improve their application profiles. I think it would be beneficial to expand the document to also target students who are not yet ready for the application process to help them improve on all criteria considered during admission. I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier Ph.D. programs. However, students from underrepresented small colleges, both in the United States and internationally, often encounter limited research opportunities within their institutions. Thankfully, there exists an alternative pathway for these aspiring researchers to gain valuable research experience, even without being physically present at a university.
